\myTitle

### 2024-03-08 ckmr-sim-evaluation.Rmd
Update the paper a bit.

# This does the simulate_Qij step and then also estimates the FPRs.
ckmr_simQ <- function(C) {
  simulate_Qij(
    C, 
    calc_relats = c("PO", "FS", "HS", "U"),
    sim_relats = c("PO", "FS", "HS", "U"),
    unlinked = FALSE,
    pedigree_list = pedigrees
  )
}

@@

Q <- ckmr_simQ(C)
```

>>>>

# This does the simulate_Qij step and then also estimates the FPRs.
ckmr_simQ <- function(C, unlinked = FALSE) {
  simulate_Qij(
    C,
    reps = 5e4,
    calc_relats = c("PO", "FS", "HS", "U"),
    sim_relats = c("PO", "FS", "HS", "U"),
    unlinked = unlinked,
    pedigree_list = pedigrees
  )
}

@@

Q_link <- ckmr_simQ(C)
Q_unl <- ckmr_simQ(C, unlinked = TRUE)

# get FPRs and FNRs for a variety of relationships.

# first for resolving unrelated individuals from PO and FS
mc_sample_simple(
  Q = Q_link,
  nu = c("PO", "FS"),
  de = "U",
  tr = "U",
  method = "IS",
  FNRs = seq(0.01, 0.3, by = 0.01)
)
```

### 2024-03-08 tex/body-text-of-paper.tex
Update the paper a bit.

Genetic markers have long been used in the 
management of Pacific salmon, and salmon conservation and management has been
a driving force behind a number of advances in molecular ecology,
from data generation techniques \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing}
to statistical methodology
\citep{smouse1990genetic,anderson2002model,pella2006gibbs}.
Two broadly applicable innovations that have been
actively fostered by the Pacific salmon research community are genetic
stock identification
(GSI:~\citealt{milner1982genetic,beacham2004dna,seeb2007development})
and parentage based tagging
(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).

@@

The ascendance over the last decade of parentage based tagging (PBT) as a management
tool for Pacific salmon adds another factor to consider when developing a GSI baseline.
Since the first proposal \citep{anderson2005description} to replace or augment the coded-wire tag
programme \citep{nandor2010overview}
with PBT, it has been noted that one of the major advantages of a genetic program for
PBT is that the genetic markers used for PBT could also be useful for GSI (and
vice versa).  This

Here, we present a reference baseline for Chinook salmon, targeted to the complex population 
structure within the Central Valley of California. Chinook salmon of the two main river basins---the 
Sacramento and the San Joaquin---within the Central Valley exhibit the greatest run-timing diversity
within the species.  With four recognized ecotypes, delineated primarily on the basis of run timing
(fall-, late-fall-, winter-, and spring-run), adult Chinook salmon can be found migrating or residing
in freshwater every month of the year in California. NEED MORE HERE. Certainly we should cite
\citet{yoshiyama1998historical}.

@@

More verbiage

@@

A number of SNP markers from earlier discovery efforts \citep{clemento2011discovery} had already 
been validated as useful for GSI
and PBT in California \citep{clemento2014evaluation}. These markers were originally assayed
using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays (CITE).
In order to type these markers using amplicon sequencing, we followed the methods in
\citet{campbell2015genotyping} and added the Illumina small RNA sequencing primer and 
read 2 sequencing primer to the SNPtype Specific Target Amplification Primer and Locus 
Specific Primers respectively.  We then dropped any loci that produced genotypes that were 
frequently inconsistent with genotypes produced with the SNPtype assays and any loci that 
did not amplify well or took over PCRs when run with the other loci in the panel.  We then 
used Freebayes \citep{garrison2012haplotype} to identify non mulltinucleotide or complex variants
in addition the the SNPs targeted by the original TaqMan assays.

@@

We undertook a new discovery process to develop additional multiallelic
microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook.
NEED TO DESCRIBE.

After we designed these loci, we compiled the reference sequences for these and the
other loci and mapped them to the Chinook v1.0 genome assembly \citep{christensen2018chinook}
with both the Burrows-Wheeler Aligner (CITE) and the BLAST-like Alignment tool \citep{kent2002blat}.
Genome locations found with both aligners were assumed to be the most likely locations
of these target amplicons in the genome.

We include amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
between spring-run and fall-run fish described in \citet{thompson2020complex}.

Need more here.

@@

@@

>>>>

Genetic markers have long been used in the 
management of Pacific salmon, and salmon conservation and management have been
behind a number of advances in molecular ecology,
from data generation techniques \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing}
to statistical methodology
\citep{smouse1990genetic,anderson2002model,pella2006gibbs}.
Two broadly applicable innovations that have been
actively fostered by the Pacific salmon research community are genetic
stock identification
(GSI:~\citealt{milner1982genetic,beacham2004dna,seeb2007development})
and parentage based tagging
(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).

@@

The ascendance over the last decade of parentage based tagging (PBT) as a management
tool for Pacific salmon adds another factor to consider when developing a GSI baseline.
Since the first proposal \citep{anderson2005description} to replace or augment the coded-wire tag
programme \citep{nandor2010overview}
with PBT, it has been noted that one of the major advantages of a genetic program for
PBT is that the genetic markers used for PBT could also be useful for GSI (and
vice versa).

Here, we present a reference baseline for Chinook salmon, targeted to the complex population 
structure within the Central Valley of California. Chinook salmon of the two main river basins---the 
Sacramento and the San Joaquin---within the Central Valley exhibit the greatest run-timing diversity
within the species.  With four recognized ecotypes, delineated primarily on the basis of run timing
(fall-, late-fall-, winter-, and spring-run), adult Chinook salmon can be found migrating or residing
in freshwater every month of the year in California. \comment{NEED MORE HERE. Certainly we should cite
\citet{yoshiyama1998historical}.}

@@

\comment{More verbiage}

@@

A number of SNP markers from earlier discovery efforts \citep{clemento2011discovery} had already 
been validated as useful for GSI
and PBT in California \citep{clemento2014evaluation}. These markers were originally assayed
using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays \comment{(CITE)}.
In order to type these markers using amplicon sequencing, we followed the methods in
\citet{campbell2015genotyping} and added the Illumina small RNA sequencing primer and 
read 2 sequencing primer to the SNPtype Specific Target Amplification Primer and Locus 
Specific Primers respectively.  We then dropped any loci that produced genotypes that were 
frequently inconsistent with genotypes produced with the SNPtype assays and any loci that 
did not amplify well or those that amplified so well that they were grossly over-represented
among the reads on the sequencer.  We then 
used Freebayes \citep{garrison2012haplotype} to identify all variants on the fragment (not just the SNPs targeted by the original TaqMan assays).  This allows us to score each locus as a
microhaplotype \citep{baetscher2018microhaplotypes} when multiple variants occur in the sequence.  We filtered out variants
that were considered multinucleotide or complex variants by Freebayes.

@@

We undertook a new discovery process to develop additional multiallelic
microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook.
\comment{NEED TO DESCRIBE. How did we look for these, etc. What constituted a
candidate to then follow up on with amplicon sequencing, etc.}

After designing and validating an amplicon-sequencing approach to type both the
existing SNPtype assays and the product of the additional microhaplotype
discovery process we identified their location with the genome. We mapped the reference
sequences for these markers 
to the Chinook v2.0 genome assembly \citep{christensen2018chinook}
with both the Burrows-Wheeler Aligner \citep{bwa-mem2009} and the BLAST-like Alignment tool \citep{kent2002blat}.
We report the genomic coordinates these target amplicons found with both alignment methods.

@@

We include amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
between spring-run and fall-run fish described in \citet{thompson2020complex}.
\comment{Need more here. I personally am of the thought that we should not include these
for GSI, but should use them for identifying the genotypes of fish assigned to populations
using the rest of the markers.}

@@

Of \comment{XX} existing SNPtype assays, we were able to convert \comment{YY}
into reliable amplicon-sequenced assays that were included in the panel.

We tested \comment{XX} candidates from our additional microhaplotype discovery and
retained \comment{185} of those in the panel.

Of these \comment{185} loci that originated from SNPtype assays or from our additional
microhaplotype discovery, \comment{179} of them mapped identically to a single
location on an assembled chromosome in the Otsh\_v2.0 genome, using both bwa mem and BLAT.
Of the remaining 6 loci, one of them was mapped to the same location by BLAT and bwa mem, but
the alignments differed by one base pair in the length of an embedded insertion, two of them had a single secondary alignment on a different chromosome
that was identical in both bwa mem and BLAT, two of them had secondary alignments that
disagreed slightly between BLAT and bwa mem, and one of them had multiple
non-primary alignments (\comment{SuppTable}).
